A 38 Million Year Old Mini-Neptune in the Kepler Field

### 2021-09-23 paper/ms.tex
update ms
Co-authored-by: adamkraus <alk@astro.as.utexas.edu>

%TODO: might want to revise this text to better frame the boil-off phase...
%TODO: better citations on radius evolution?
For instance, the ``mini-Neptunes'', thought to be made of molten
rocky cores \citep{kite_atmosphere_2020} and extended atmospheric
envelopes of hydrogen and helium, are expected to shrink in size by
factors of several over their first $10^8$ years.  Specifically, in
the models of \citet{owen_atmospheres_2016} and
\citet{owen_constraining_2020}, the $\approx5\,M_\oplus$ planets start
with sizes of 4--12\,$R_\oplus$ shortly after the time of disk
dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  While the majority of this change
is expected to occur within the first few million years of disk
dispersal \citep{owen_atmospheres_2016}, the combination of stellar
irradiation and internal heat may also power a more gradual outflow
which can eventually deplete or entirely strip the envelope from the
rocky core \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
Discovering young planets, measuring their masses, and detecting their
atmospheric outflows are key steps toward testing this paradigm, which
is often invoked to explain the observed radius distribution of mature
exoplanets \citep{Fulton_et_al_2017}.

@@

Beyond the kinematic and rotational lines of evidence for the youth of
\sn, there is a final consistency check: the abundance of
Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet \citep[{\it
e.g.},][]{soderblom_ages_2014}.  We acquired an iodine-free spectrum
from Keck/HIRES on the night of 2021 March 26 using the standard setup
and reduction techniques of the California Planet Survey
\citep{howard_cps_2010}.  Following the equivalent width measurement
procedure described by \citet{bouma_2021_ngc2516}, we find ${\rm
EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   %Note that this value does not
%correct for the Fe\,\textsc{I} blend at 6707.44\AA. 
Given the effective temperature of \sn\ (Table~\ref{tab:starparams}),
this measurement is also in line with expectations for a 38\,Myr star
({\it e.g.}, \citealt{randich_gaiaeso_2018};
\citealt{bouma_cluster_2020}).

@@

		\subfloat{
			\includegraphics[width=0.5\textwidth]{f5a.pdf}
			\includegraphics[width=0.5\textwidth]{f5b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The light curve of Kepler\,1627.}
    {\it Top}: 
    The Kepler data span 1{,}437 days (3.9 years), sampled at
    30 minute cadence;  a 100 day segment is shown.  The
    top panel shows the \texttt{PDCSAP} median-subtracted flux in
    units of parts-per-thousand ($\times 10^{-3}$).  The dominant
    signal is induced by starspots.  The stellar
    variability model (orange line) is subtracted below, revealing the
    transits of \pn.  The
    online Figure Set spans the entire 3.9 years of observations.
    {\it Bottom}:
    Phase-folded transit of Kepler 1627Ab with stellar variability
    removed.  Windows over 20 hours ({\it left}) and the entire orbit
    ({\it right}) are shown.  The 2-$\sigma$ model uncertainties and
    the best-fit model are the light purple band and the dark purple
    line.  Gray points are individual flux measurements; black points
    bin these to 20 minute intervals, and have a representative
    1-$\sigma$ error bar in the lower right of each panel.  The
    asymmetric residual during transit appears to be robust to
    different detrending methods.
    \label{fig:lc}
  }
\end{figure*}

@@

The second line of confirmatory evidence comes by analyzing the
individual Kepler transit times. We isolated each of the 144 observed
transits to within $\pm4.5$\,hr of each transit, and fitted each
window with both {\it i)} a local second-order polynomial and transit,
and {\it ii)} a local linear trend.  We let the mid-time of each
transit float, and then calculated the residual between the measured
mid-time and that of a periodic orbit.  This residual, the transit
timing variation (TTV), is plotted in Figure~\ref{fig:obliquity}
against the local linear slope.  A significant correlation of $-0.0564
\pm 0.0098\ {\rm ppt\,day}^{-1}$ is observed.  Fewer than ten Kepler
Objects of Interest have shown this correlation
\citep{holczer_time_2015}, which is most readily interpreted as a TTV
induced by unresolved starspot crossings \citep{mazeh_time_2015}.
This is only possible if the planet transits the primary star, which
excludes a background eclipsing binary scenario, and therefore
confirms that \pn\ is a planet.  It also implies that the planet's
orbit is prograde.  The latter point assumes that the dominant
photometric variability is induced by dark spots, and not bright
faculae.  Given the observed transition of Sun-like stellar
variability from spot to faculae-dominated regimes, we expect this
assumption to be reasonably secure
\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.

@@

Kepler\,1627Ab provides a new extremum in the ages of the Kepler
planets, and opens multiple avenues for further study.  Further
observations of spectroscopic transits could yield a quantification of
the stellar obliquity, and would confirm or refute the expected
prograde orbital geometry.  Similarly, transit spectroscopy aimed at
detecting atmospheric outflows could yield insight into the
evolutionary state of the atmosphere \citep[{\it
e.g.},][]{ehrenreich_giant_2015,spake_helium_2018,vissapragada_2020}.
Observations aimed at quantifying the amount of high-energy
irradiation incident on the planet would also help in clarifying the
expected outflow rate \citep[{\it e.g.},][]{poppenhaeger_2021}.
Finally, a challenging but informative quantity to measure would be
the planet's mass.  This might be possible through long-term
monitoring of the transit times and durations, or through a
high-cadence multi-color radial velocity campaign.  Measured at
sufficient precision, the mass could yield constraints on both the
planet's composition and its initial entropy
\citep{owen_constraining_2020}.

Separately, the Kepler data may yet contain additional information.
For instance, one possible explanation for the transit asymmetry shown
in Figure~\ref{fig:lc} is that of a dusty asymmetric outflow.  Other
possible explanations such as favored longitudes for starspot
crossings seem strained, for reasons discussed in
Appendix~\ref{app:gptransit}.  Dusty outflows are theoretically
expected for young mini-Neptunes, and the amplitude of the observed
asymmetry is also roughly consistent with predictions
\citep{wang_dai_2019}.  Beyond the planetary transits,
Appendix~\ref{app:flare} highlights a separate odd coincidence in the
Kepler data, in the arrival time distribution of stellar flares.  We
encourage its exploration by investigators more versed in the topic
than ourselves.

@@

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f17.pdf}
			\includegraphics[width=0.48\textwidth]{f6.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Kepler 1627 is a binary.} {\it Left:} High-resolution imaging
    from Gemini-North/`Alopeke and Keck/NIRC2 shows an $\approx$M2.5V
    companion at $\rho \approx 0\farcs16$, which corresponds to a
    projected separation of $53\pm4$\,AU.  The inset shows a cutout of
    the NIRC2 image (North is up, East is left, scale is set by the
    separation of the binary).  The lines show 5-$\sigma$ contrast
    limits for the `Alopeke filters, and 6-$\sigma$ contrast limits
    for NIRC2.  {\it Left:} Gaia EDR3 renormalized
    unit weight error (RUWE) point estimates for Kepler\,1627A and
    other members of the $\delta$\,Lyr cluster reported by
    \citet{KounkelCovey2019}.  Since other members of the cluster with
    similar brightnesses have comparable degrees of photometric
    variability, the high RUWE suggests that Kepler\,1627 is a
    binary. 
    \label{fig:kep1627binary}
	}
\end{figure*}

@@

Figure~\ref{fig:kep1627binary} summarizes the results of the
high-resolution imaging.  The Gaia EDR3 parallax for the primary
implies a projected separation of $53 \pm 4$\,AU, assuming the
companion is bound.  Although the companion is unresolved in the Gaia
source catalog (there are no candidates brighter than $G < 20.5$ mag
within $\rho < 120\arcsec$), its existence was also suggested by the
primary star's large RUWE, relative to other members of the \cn.
Based on the apparent separation, the binary orbital period is of
order hundreds of years.  The large RUWE is therefore more likely to
be caused by a PSF-mismatch skewing the Gaia centroiding during
successive scans, rather than true astrometric motion.  Regardless,
given the large observed RUWE of the primary, and the low geometric
probability that a companion imaged at $\rho \approx 0\farcs16$ is a
chance line-of-sight companion, we proceed under the assumption that
the companion is bound, and that Kepler\,1627 is a binary.  Given the
distance and age, the models of \citet{baraffe_new_2015} imply a
companion mass of $M_{\rm B} \approx 0.33 M_{\odot}$ and companion
temperature of $T_{\rm eff,B} \approx 3450$ K.  The corresponding
spectral type is roughly M2.5V \citep{pecaut_mamajek_2013}.
% \footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt}, version \texttt{2021/03/02}.}

>>>>

%TODO: might want to revise this text to better frame the boil-off phase...
%TODO: better citations on radius evolution?
For instance, the ``mini-Neptunes'', thought to be made of molten
rocky cores \citep{kite_atmosphere_2020} and extended atmospheric
envelopes of hydrogen and helium, are expected to shrink in size by
factors of several over their first $10^8$ years.  Specifically, in
the models of \citet{owen_atmospheres_2016} and
\citet{owen_constraining_2020}, the $\approx5\,M_\oplus$ planets start
with sizes of 4--12\,$R_\oplus$ shortly after the time of disk
dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  While the majority of this change
is expected to occur within the first few million years after disk
dispersal \citep{owen_atmospheres_2016}, the combination of stellar
irradiation and internal heat may also power a more gradual outflow
which can eventually deplete or entirely strip the envelope from the
rocky core \citep{Owen_Wu_2013,ginzburg_corepowered_2018}.
Discovering young planets, measuring their masses, and detecting their
atmospheric outflows are key steps toward testing this paradigm, which
is often invoked to explain the observed radius distribution of mature
exoplanets \citep{Fulton_et_al_2017}.

@@

Beyond the kinematic and rotational lines of evidence for the youth of
\sn, there is a final consistency check: the abundance of
Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet \citep[{\it
e.g.},][]{soderblom_ages_2014}.  We acquired an iodine-free spectrum
from Keck/HIRES on the night of 2021 March 26 using the standard setup
and reduction techniques of the California Planet Survey
\citep{howard_cps_2010}.  Following the equivalent width measurement
procedure described by \citet{bouma_2021_ngc2516}, we find ${\rm
EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   %Note that this value does not
%correct for the Fe\,\textsc{I} blend at 6707.44\AA. 
Given the effective temperature of \sn\ (Table~\ref{tab:starparams}),
this measurement is in agreement with expectations for a 38\,Myr star
({\it e.g.}, as measured in IC~2602 by
\citealt{randich_gaiaeso_2018}).  It is also larger than any ${\rm
EW}_{\rm Li}$ found by \citet{berger_identifying_2018} in their
analysis of 1{,}301 Kepler spectra.

@@

		\subfloat{
			\includegraphics[width=0.5\textwidth]{f5a.pdf}
			\includegraphics[width=0.5\textwidth]{f5b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The light curve of Kepler\,1627.}
    {\it Top}: 
    The Kepler data span 1{,}437 days (3.9 years), sampled at
    30 minute cadence;  a 100 day segment is shown.  The
    top panel shows the \texttt{PDCSAP} median-subtracted flux in
    units of parts-per-thousand ($\times 10^{-3}$).  The dominant
    signal is induced by starspots.  The stellar
    variability model (orange line) is subtracted below, revealing the
    transits of \pn.  The
    online Figure Set spans the entire 3.9 years of observations.
    {\it Bottom}:
    Phase-folded transit of Kepler 1627Ab with stellar variability
    removed.  Windows over 20 hours ({\it left}) and the entire orbit
    ({\it right}) are shown, and the residual after subtracting the
    transit in the bottom-most row.  The 2-$\sigma$ model uncertainties
    and the best-fit model are the light purple band and the dark
    purple line.  Gray points are individual flux measurements; black
    points bin these to 20 minute intervals, and have a representative
    1-$\sigma$ error bar in the lower right of each panel.  The
    asymmetric residual during transit appears to be robust to
    different detrending methods.
    \label{fig:lc}
  }
\end{figure*}

@@

The second line of confirmatory evidence comes by analyzing the
individual Kepler transit times. We isolated each of the 144 observed
transits to within $\pm4.5$\,hr of each transit, and fitted each
window with both {\it i)} a local second-order polynomial and transit,
and {\it ii)} a local linear trend.  We let the mid-time of each
transit float, and then calculated the residual between the measured
mid-time and that of a periodic orbit.  This residual, the transit
timing variation (TTV), is plotted in Figure~\ref{fig:obliquity}
against the local linear slope.  A significant correlation of $-0.0564
\pm 0.0098\ {\rm ppt\,day}^{-1}$ is observed.  Fewer than ten Kepler
Objects of Interest have shown this correlation
\citep{holczer_time_2015}, which is most readily interpreted as a TTV
induced by unresolved starspot crossings \citep{mazeh_time_2015}.
This is only possible if the planet transits the primary star, which
excludes a background eclipsing binary scenario, and therefore
confirms that \pn\ is a planet.  It also implies that the planet's
orbit is prograde.  The latter point assumes that the dominant
photometric variability is induced by dark spots, and not bright
faculae.  Given the observed transition of Sun-like stellar
variability from spot to faculae-dominated regimes between young and
old ages, we expect this assumption to be reasonably secure
\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.

@@

Kepler\,1627Ab provides a new extremum in the ages of the Kepler
planets, and opens multiple avenues for further study.  Further
observations of spectroscopic transits could yield a quantification of
the stellar obliquity, and would confirm or refute the expected
prograde orbital geometry.  Similarly, transit spectroscopy aimed at
detecting atmospheric outflows could yield insight into the
evolutionary state of the atmosphere \citep[{\it
e.g.},][]{ehrenreich_giant_2015,spake_helium_2018,vissapragada_2020}.
Observations aimed at quantifying the amount of high-energy
irradiation incident on the planet would also help in clarifying the
expected outflow rate \citep[{\it e.g.},][]{poppenhaeger_2021}.
Finally, a challenging but informative quantity to measure would be
the planet's mass.  This might be possible through a high-cadence
multi-color radial velocity campaign, or through long-term monitoring
of the transit times and durations.  Measured at sufficient precision,
the mass could yield constraints on both the planet's composition and
its initial entropy \citep{owen_constraining_2020}.

Separately, the Kepler data may yet contain additional information.
For instance, one possible explanation for the transit asymmetry shown
in Figure~\ref{fig:lc} is that of a dusty asymmetric outflow.  Other
possible explanations such as favored longitudes for starspot
crossings seem strained, for reasons discussed in
Appendix~\ref{app:gptransit}.  Dusty outflows are theoretically
expected for young mini-Neptunes, and the amplitude of the observed
asymmetry is also roughly consistent with predictions
\citep{wang_dai_2019}.  Beyond the planetary transits,
Appendix~\ref{app:flare} highlights a separate oddity in the Kepler
data, in the arrival time distribution of stellar flares.  We
encourage its exploration by investigators more versed in the topic
than ourselves.

@@

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f17.pdf}
			\includegraphics[width=0.48\textwidth]{f6.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Kepler 1627 is a binary.} {\it Left:} High-resolution imaging
    from Gemini-North/`Alopeke and Keck/NIRC2 shows an $\approx$M2.5V
    companion at $\rho \approx 0\farcs16$, which corresponds to a
    projected separation of $53\pm4$\,AU.  The inset shows a cutout of
    the stacked NIRC2 image (North is up, East is left, scale is set by the
    separation of the binary).  The lines show 5-$\sigma$ contrast
    limits for the `Alopeke filters, and 6-$\sigma$ contrast limits
    for NIRC2.  Inside the diffraction limit (estimated as the gray
    band), the detection limits are less reliable.  {\it Left:} Gaia
    EDR3 renormalized unit weight error (RUWE) point estimates for
    Kepler\,1627A and other members of the $\delta$\,Lyr cluster
    reported by \citet{KounkelCovey2019}.  Since other members of the
    cluster with similar brightnesses have comparable degrees of
    photometric variability, the high RUWE suggests that Kepler\,1627
    is a binary. 
    \label{fig:kep1627binary}
	}
\end{figure*}

@@

Figure~\ref{fig:kep1627binary} summarizes the results of the
high-resolution imaging.  The Gaia EDR3 parallax for the primary
implies a projected separation of $53 \pm 4$\,AU, assuming the
companion is bound.  Although the companion is unresolved in the Gaia
source catalog (there are no comoving, codistant candidate companions brighter than $G < 20.5$ mag
within $\rho < 120\arcsec$), its existence was also suggested by the
primary star's large RUWE, relative to other members of the \cn.
Based on the apparent separation, the binary orbital period is of
order hundreds of years.  The large RUWE is therefore more likely to
be caused by a PSF-mismatch skewing the Gaia centroiding during
successive scans, rather than true astrometric motion.  Regardless,
given the large observed RUWE of the primary, and the low geometric
probability that a companion imaged at $\rho \approx 0\farcs16$ is a
chance line-of-sight companion, we proceed under the assumption that
the companion is bound, and that Kepler\,1627 is a binary.  Given the
distance and age, the models of \citet{baraffe_new_2015} imply a
companion mass of $M_{\rm B} \approx 0.33 M_{\odot}$ and companion
temperature of $T_{\rm eff,B} \approx 3450$ K.  The corresponding
spectral type is roughly M2.5V \citep{pecaut_mamajek_2013}.
% \footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt}, version \texttt{2021/03/02}.}
